Automated Formal Synthesis of Provably Safe Digital Controllers for Continuous Plants

### 2019-01-25 automatica/paper.tex
sections

\subsection{Interval arithmetic over errors in numeric representation} 
We use finite-precision arithmetic to model the plant. 
This is an approximation that speeds up each CEGIS iteration, 
however it necessitates a further stage where we verify that the errors introduced by the approximation have not resulted in a controller that is unsafe when executed on a plant expressed over real numbers.  
In this stage, we represent the plant using double-precision floating-point
and we use the Boost interval arithmetic library~\cite{DBLP:journals/tcs/BronnimannMP06} to bound the error in this representation. 
We employ a compositional numeric library to model the fixed point arithmetic for the controller~\cite{CNL} within double-precision floating-point numbers. 
We check that the controller is safe starting from each vertex of the set of initial states,  
and show that this is sufficient to prove safety from any state in this set.

@@

The fixed-point arithmetic of the digital controller is performed on the upper and lower bound of the intervals independently and the upper and lower bound of the result taken as the interval result\note{Would an example make this clearer?, e.g., consider the conversion from plant precision to controller precision performed by the ADC on a single state value. The state value is represented as an interval $\{x.high, x.low\}$, and the result of the conversion is an interval where the upper bound is the conversion of $x.high$ and the lower bound is the conversion of $x.low$?}. 
Since the precision of the floating point domain is greater than the precision of the controller, this is guaranteed to bound the real behaviour of the controller.

@@

An additional error is due to the representation of the plant dynamics, namely 
%
{\scriptsize
\begin{equation*}
x_{k+1} =\mathcal{F}_{\langle I_p,F_p \rangle}(A_d) \mathcal{F}_{\langle I_p,F_p \rangle}(x_{k}) + \mathcal{F}_{\langle I_p,F_p \rangle}(B_d)\mathcal{F}_{\langle I_p,F_p \rangle}(u_{k}).
\end{equation*}
}
We encompass this error using interval arithmetic~\cite{moore1966interval} in the verification phase \note{please add more details here}.

\subsubsection{Stability of closed-loop models with fixed-point controller error}
%-------------------------------
\label{sec:stab_FWL}
The validity of Jury's criterion~\cite{fadali} relies on the representation of the closed-loop dynamics $x_{k+1} = (A_d - B_dK) x_k$ at infinite precision.  
When we employ a fixed-point arithmetic digital controller, the operation above can be expressed as follows, where we use $\mathcal{F}_{\langle I_c,F_c \rangle}$ preceding a variable to indicate that variable is converted into the fixed-point precision given by $\mathcal{F}_{\langle I_c,F_c \rangle}$: 
%
\begin{align*}
x_{k+1} &= A_d \cdot x_{k} -B_d(\mathcal{F}_{\langle I_c,F_c \rangle}(K)\cdot\mathcal{F}_{\langle I_c,F_c \rangle}(x_{k})).  \\
x_{k+1} &= (A_d  - B_dK) \cdot x_k + B_dK\delta, 
%\\ x_{k+1}&=A_d x_{k} + B_du_{k}
\end{align*}
%
where $\delta$ is the maximum error that can be introduced by the digital controller in one step, 
i.e., by reading the states values once and multiplying by $K$ once.  
We derive the closed form expression for $x_n$ recursively, as follows:
%
\small
\begin{align*}
x_{1} &= (A_d  - B_dK)x_0 + B_dK\delta \\
x_{2} %&= (A_d  - B_dK)((A_d  - B_dK)x_0 + B_dK\delta ) + B_dK\delta \\
 &=(A_d  - B_dK)^2x_0 + (A_d  - B_dK)B_dK\delta + B_dK\delta \\
%x_{3}  &=(A_d  - B_dK)^3x_0 + (A_d  - B_dK)^2B_dK\delta + (A_d  - B_dK)B_dK\delta + B_dK\delta\\
x_{n} &= (A_d  - B_dK)^nx_0 + (A_d  - B_dK)^{n-1}B_dK\delta +\\
&\phantom{{}=1} ... + (A_d  - B_dK)^1B_dK \delta + B_dK\delta \\
  &= (A_d - B_dK)^nx_0 + \sum_{i=0}^{i=n-1}(A_d - B_dK)^iB_dk\delta. 
\end{align*}
\normalsize
%The definition of asymptotic stability is that the system converges to a
%reference signal, in this case we use no reference signal so 
Recall that a closed-loop asymptotically stable system will converge to the origin.  
We know that the original system with an infinite-precision controller is stable, 
because we have synthesized it to meet Jury's criterion.  
Hence, $(A_d - B_dK)^n x_0$  must converge to zero as $n \uparrow \infty$. 
%T%o show that a fixed-point implementation of the same controller will
%converge to a finite value, we need to show that the infinite sum converges. 
Furthermore, the power series of a square matrix $T$ converges~\cite{horn1990matrix} iff the eigenvalues of the matrix are less
than~1, and the limit results in 
%
%\begin{align*}
$\sum_{i=0}^{\infty}T^i  = (I - T)^{-1}$, 
%\end{align*}
%
where $I$ is the identity matrix. 
Thus, the closed-loop model converges to the value 
%
\begin{align*}
0 + (I - A_d + B_dK)^{-1}B_dk\delta \,. 
\end{align*}
%
As a result, if the value $(I - A_d + B_dK)^{-1}B_dk\delta$ is within the safe set of states given by the safety specification
then the synthesized fixed-point controller results in a safe
closed-loop model.  The convergence to a finite value, however, will not
make it asymptomatically stable.
However,
since in this paper we require stability only as a precursor to safety, it
is sufficient to check that the perturbed model converges to a
neighborhood of the equilibrium within the safe set.

@@

\note{Please turn itemisation below into sub(sub?)sections.}

@@

%\todo{please eventually itemize next}
\begin{enumerate}

\item The inductive synthesis phase ({\sc synthesize}) uses BMC to compute a
candidate solution $K$ that satisfies both the stability requirement and the safety specification, 
within a finite precision model.  
In order to synthesize a controller that satisfies the stability requirement, 
we need the computed characteristic polynomial to satisfy Jury's criterion~\cite{fadali} (see Section~\ref{ssec:stabspecification}).

@@

\item The first {\sc verify} stage, {\sc safety} is shown in Alg.~\ref{alg:safetycheck}. 
The algorithm checks that the candidate
solution $K$, which we have synthesized to be safe for at least one initial
state, is safe for \emph{all} possible initial states, i.e., it does not reach
the (complement) unsafe state within $k$ steps.  After unfolding the transition system corresponding
to the previously synthesized controller $k$ steps, we check that the safety
specification holds for any initial state.

@@

\item The second {\sc verify} stage, {\sc precision}, 
 restores soundness with respect to the plant precision
by using interval arithmetic \cite{moore1966interval} to validate the 
operations performed by the previous stage.

\item The third {\sc verify} stage, {\sc complete}, checks that the current
$k$ is large enough to ensure safety for any further time steps.  
Here, we compute the completeness threshold $\overline{k}$ for the current candidate controller $K$ and
check that $k{\geq}\overline{k}$.  This is done by computing the number of time steps
required for the states to have completed a 360\textdegree circle, as illustrated in Fig.~\ref{fig:ct}.

@@

%-------------------------------
\section{DSSynth: An Automated Digital Controller Synthesis Tool for Physical Plants}
\label{sec:dssynthtool}
%-------------------------------

>>>>

\subsection{Interval arithmetic over errors in numeric representation} 
We use finite-precision arithmetic to model the plant. 
This is an approximation that speeds up each CEGIS iteration, 
however it necessitates a further stage where we verify that the errors introduced by the approximation have not resulted in a controller that is unsafe when executed on a plant expressed over real numbers.  
In this stage, we represent the plant using double-precision floating-point
and we use the Boost interval arithmetic library~\cite{DBLP:journals/tcs/BronnimannMP06} to bound the error in this representation. 
We employ a compositional numerical library to model the fixed point arithmetic for the controller~\cite{CNL} within double-precision floating-point numbers. 
We check that the controller is safe starting from each vertex of the set of initial states,  
and show that this is sufficient to prove safety from any state in this set. \note{add forward reference to proof.}

@@

The fixed-point arithmetic of the digital controller is performed on the upper and lower bound of the intervals from above independently, [I DO NOT UNDERSTAND THE REMAINING PART OF THE SENTENCE.] and the upper and lower bound of the result taken as the interval result\note{Would an example make this clearer?, e.g., consider the conversion from plant precision to controller precision performed by the ADC on a single state value. The state value is represented as an interval $\{x.high, x.low\}$, and the result of the conversion is an interval where the upper bound is the conversion of $x.high$ and the lower bound is the conversion of $x.low$?}. 
Since the precision of the floating point domain is greater than the precision of the controller, this is guaranteed to bound the real behaviour of the controller.

@@

An additional error is due to the representation of the plant dynamics, namely 
%
{\scriptsize
\begin{equation*}
x_{k+1} =\mathcal{F}_{\langle I_p,F_p \rangle}(A_d) \mathcal{F}_{\langle I_p,F_p \rangle}(x_{k}) + \mathcal{F}_{\langle I_p,F_p \rangle}(B_d)\mathcal{F}_{\langle I_p,F_p \rangle}(u_{k}).
\end{equation*}
}
We encompass this error using interval arithmetic~\cite{moore1966interval} in the verification phase \note{please add more details here on the ``verification phase''}.

\subsubsection{Stability of closed-loop models with fixed-point controller error}
%-------------------------------
\label{sec:stab_FWL}
The validity of Jury's criterion~\cite{fadali} relies on the representation of the closed-loop dynamics $x_{k+1} = (A_d - B_dK) x_k$ at infinite precision.  
When we employ a fixed-point arithmetic digital controller, the operation above can be expressed as follows, where we use $\mathcal{F}_{\langle I_c,F_c \rangle}$ preceding a variable to indicate that variable is converted into the fixed-point precision given by $\mathcal{F}_{\langle I_c,F_c \rangle}$: 
%
\begin{align*}
x_{k+1} &= A_d \cdot x_{k} -B_d(\mathcal{F}_{\langle I_c,F_c \rangle}(K)\cdot\mathcal{F}_{\langle I_c,F_c \rangle}(x_{k})).  
\end{align*}
This translates to 
\begin{align*}
x_{k+1} &= (A_d  - B_dK) \cdot x_k + B_dK\delta, 
%\\ x_{k+1}&=A_d x_{k} + B_du_{k}
\end{align*}
%
where $\delta$ is the maximum error that can be introduced by the digital controller in one step, 
i.e., by reading the states values once and multiplying by $K$ once.  
We derive the closed form expression for $x_n$ recursively, as follows:
%
\small
\begin{align*}
x_{1} &= (A_d  - B_dK)x_0 + B_dK\delta \\
x_{2} %&= (A_d  - B_dK)((A_d  - B_dK)x_0 + B_dK\delta ) + B_dK\delta \\
 &=(A_d  - B_dK)^2x_0 + (A_d  - B_dK)B_dK\delta + B_dK\delta \\
%x_{3}  &=(A_d  - B_dK)^3x_0 + (A_d  - B_dK)^2B_dK\delta + (A_d  - B_dK)B_dK\delta + B_dK\delta\\
x_{n} &= (A_d  - B_dK)^nx_0 + (A_d  - B_dK)^{n-1}B_dK\delta +\\
&\phantom{{}=1} ... + (A_d  - B_dK)^1B_dK \delta + B_dK\delta \\
  &= (A_d - B_dK)^nx_0 + \sum_{i=0}^{i=n-1}(A_d - B_dK)^iB_dk\delta. 
\end{align*}
\normalsize
%The definition of asymptotic stability is that the system converges to a
%reference signal, in this case we use no reference signal so 
Recall that a closed-loop asymptotically stable system will converge to the origin.  
We know that the original system with an infinite-precision controller is stable, 
because we have synthesized it to meet Jury's criterion.  
Hence, $(A_d - B_dK)^n x_0$  must converge to zero as $n \uparrow \infty$. 
%T%o show that a fixed-point implementation of the same controller will
%converge to a finite value, we need to show that the infinite sum converges. 
Furthermore, the power series of a square matrix $T$ converges~\cite{horn1990matrix} iff the eigenvalues of the matrix are less
than~1, and the limit results in 
%
%\begin{align*}
$\sum_{i=0}^{\infty}T^i  = (I - T)^{-1}$, 
%\end{align*}
%
where $I$ is the identity matrix. 
Thus, the closed-loop model converges to the value 
%
\begin{align*}
0 + (I - A_d + B_dK)^{-1}B_dk\delta \,. 
\end{align*}
%
As a result, if the value $(I - A_d + B_dK)^{-1}B_dk\delta$ is within the safe set of states given by the safety specification
then the synthesized fixed-point controller results in a safe
closed-loop model.  The convergence to a finite value, however, will not
make it asymptomatically stable.
However,
since in this paper we require stability only as a precursor to safety, it
is sufficient to check that the perturbed model converges to a
neighborhood of the equilibrium within the safe set.

@@

@@

%\todo{please eventually itemize next}
\subsubsection{Block 1}

The inductive synthesis phase ({\sc synthesize}) uses BMC to compute a
candidate solution $K$ that satisfies both the stability requirement and the safety specification, 
within a finite precision model.  
In order to synthesize a controller that satisfies the stability requirement, 
we need the computed characteristic polynomial to satisfy Jury's criterion~\cite{fadali} (see Section~\ref{ssec:stabspecification}).

@@

\subsubsection{Block 2}

The first {\sc verify} stage, {\sc safety} is shown in Alg.~\ref{alg:safetycheck}. 
The algorithm checks that the candidate
solution $K$, which we have synthesized to be safe for at least one initial
state, is safe for \emph{all} possible initial states, i.e., it does not reach
the (complement) unsafe state within $k$ steps.  After unfolding the transition system corresponding
to the previously synthesized controller $k$ steps, we check that the safety
specification holds for any initial state.

@@

\subsubsection{Block 3}

The second {\sc verify} stage, {\sc precision}, 
 restores soundness with respect to the plant precision
by using interval arithmetic \cite{moore1966interval} to validate the 
operations performed by the previous stage.

\subsubsection{Block 4}

The third {\sc verify} stage, {\sc complete}, checks that the current
$k$ is large enough to ensure safety for any further time steps.  
Here, we compute the completeness threshold $\overline{k}$ for the current candidate controller $K$ and
check that $k{\geq}\overline{k}$.  This is done by computing the number of time steps
required for the states to have completed a 360\textdegree circle, as illustrated in Fig.~\ref{fig:ct}.

@@

%-------------------------------
\section{DSSynth: A Software Tool for Automated Digital Controller Synthesis over Physical Plants}
\label{sec:dssynthtool}
%-------------------------------
